---
title: "vaccine-mini-project"
author: "Andrea Sama (A59010582)"
date: "11/28/2021"
output:
  pdf_document: default
  html_document: default
---

Comparing vaccination rates from around San Diego.

```{r}
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")

tail(vax)
```
Question 1: The column "persons_fully_vaccinated" details the number of people fully vaccinated

Question 2: The column "zip_code_tabulation_area" details the zip code tabulation area.

Question 3: The earliest date in this data set is January 5th 2021

Question 4: The latest date in this data set is November 23rd 2021/

```{r}
#skimr::skim(vax)
```

Q5: There are nine numeric columns in this data set. 

Q6: There are 8355 NA values in the persons_fully_vaccinated column.

Q7: The percent of persons_fully_vaccinated that are missing is about 10%

```{r}
#install.packages("lubridate")
library(lubridate)
```

```{r}
today()
```

```{r}
# Specify that we are using the year-month-day format
vax$as_of_date <- ymd(vax$as_of_date)
```


```{r}
today() - vax$as_of_date[1]
```
How much time since the first vaccination reported.

```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]

```

```{r}
vax$as_of_date[82908] - vax$as_of_date[1]
```

```{r}
today() - vax$as_of_date[nrow(vax)]
```

5 days have passed.


```{r}
(vax$as_of_date[nrow(vax)]-vax$as_of_date[1])/7
```
46 unique dates in the dataset.

#Working with ZIP codes

```{r}
#install.packages("zipcodeR")
library(zipcodeR)
```

```{r}
geocode_zip('92037')
```

```{r}
zip_distance('92037','92109')
```

```{r}
reverse_zipcode(c('92037', "92109") )

```

```{r}
# Pull data for all ZIP codes in the dataset
zipdata <- reverse_zipcode( vax$zip_code_tabulation_area )
zipdata

```
```{r}
# Subset to San Diego county only areas
library(dplyr)

sd <- filter(vax, county == "San Diego")

nrow(sd)
```


```{r}
unique(sd$zip_code_tabulation_area)
```

```{r}
length(unique(sd$zip_code_tabulation_area))
```


There are 107 unique zip codes in San Diego County.


```{r}
which.max(sd$age12_plus_population)

sd[60,]
```

The 92154 zip code has the highest age 12+ population.


```{r}
library(dplyr)

sd.date <- filter(vax, county == "San Diego", as_of_date == "2021-11-16")

head(sd.date)



```
```{r}
mean(sd.date$percent_of_population_fully_vaccinated, na.rm=TRUE)
```

```{r}
sd.date.na <-na.omit(sd.date)
head(sd.date.na)

mean(sd.date.na$percent_of_population_fully_vaccinated)
```



```{r}
library(ggplot2)

ggplot(sd.date, aes(sd.date$percent_of_population_fully_vaccinated, )) + geom_histogram()



```


#Focus on UCSD/La Jolla


```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```


```{r}
ggplot(ucsd) +
  aes(ucsd$as_of_date,
      ucsd$percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated")
```
```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2021-11-16")

head(vax.36)
```

```{r}
mean(vax.36$percent_of_population_fully_vaccinated)
```

```{r}
ggplot(ucsd) +
  aes(ucsd$as_of_date,
      ucsd$percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated")+
  geom_hline(yintercept=0.6640413, color="red", linetype= "dashed")
```

```{r}
fivenum(vax.36$percent_of_population_fully_vaccinated)
mean(vax.36$percent_of_population_fully_vaccinated)
```

```{r}
ggplot(vax.36) +
  aes(vax.36$percent_of_population_fully_vaccinated) +
  geom_histogram() 
```

```{r}
#vax %>% filter(as_of_date == "2021-11-16") %>%  
  #filter(zip_code_tabulation_area=="92040") %>%
  #select(percent_of_population_fully_vaccinated)
```

Below the average value.

```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)


ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  ylim(0, 1.00) +
  labs(x="Date", y="Percent Vaccinated",
       title="Vaccination rate across California",
       subtitle="only areas with a population above 36k are shown") +
  geom_hline(yintercept = 0.6640413, linetype= "dashed")
```

I think I would prefer for next weeks classes to be online, especially with the Omicron variant surge. 







