---
title: 'Class 6: R function'
author: "Andrea Sama (A59010582)"
date: "10/15/2021"
output: pdf_document
---

## Quick Rmarkdown intro

We can write text of course just like any file. We 
can style text to be **bold** or *italic*.

Do:

- this
- that
- and another thing

---

We can include some code:

```{r}
plot(1:10)
```

```{r}
#This is a comment and will not be passed to R
plot(cars)
```


## Time to write a function!

>Q1. Write a function grade() to determine an overall grade from a vector of student homework
assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an
NA value) this can be used as a score to be potentially dropped. Your final function should be
adquately explained with code comments and be able to work on an example class gradebook
such as this one in CSV format: “https://tinyurl.com/gradeinput” [3pts]

>**Hints**:  
Once you have a working function for vector inputs (such as the student1, student2, and student3 vectors
below) you can use the apply() function to work with data frame inputs such as those obtained from
read.csv().
Additional functions you will want to explore include mean(), is.na(), which.min(), which.max(), sum(),
and cor(). Remember, you can ask for help on any function by typing a question mark before the function
name e.g. ?sum.
We will walk through many of these steps together in this weeks live screen-cast. However, attempting on
your own before then is highly recommended and will be a big help for following the screen-cast.



```{r}
# Example input vectors to start with
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)
```

```{r}
min(student1)
```
First I want to find the lowest score. I can use the **min()** function to find what it is,
and the **which.min()** to find where it is (the position in the vector) 
```{r}
which.min(student1)
```
get everything but the lowest score (use - in front of which.min function)
```{r}
student1[-which.min(student1)]
```
Now I can use the **mean()** function to get the average.
```{r}
mean(student1[-which.min(student1)])
```
Does this code work for student 2?
```{r}
mean(student2[-which.min(student2)])

```
No!
Why not? Does which.min work?
```{r}
student2
which.min(student2)
```

```{r}
mean(student2)
```

```{r}
mean(student2, na.rm=TRUE)
```

```{r}
student3
```

One idea is to replace NA values with zero. Let's try it:
Step 1: find the NAs in the vectors:(from stackexchange)

```{r}
which(is.na(student2))
is.na(student2)
```
**is.na** prints out a logical answer to where the NA values are in a vector
```{r}
student2[is.na(student2)]
```

Lets replace NAs with zero
```{r}
student.prime<- student2
student.prime[is.na(student.prime)]=0
student.prime
```
Will this work now for student 2?
```{r}
student.prime<- student2
student.prime[is.na(student.prime)]=0
mean(student.prime[-which.min(student.prime)])
```
Yay! It works!

Will this work for student 3?
```{r}
student.prime<-student3
student.prime[is.na(student.prime)]=0
mean(student.prime[-which.min(student.prime)])
```
Check that this works:
```{r}
student3
```
```{r}
mean(c(90, 0, 0, 0, 0, 0, 0))
```
Yes! We did it! This works, lets simplify it!
change student.prime to x
```{r}
x<-student3
x[is.na(x)]=0
mean(x[-which.min(x)])
```

New student! with a mistake in the code!
```{r}
student4 <-c(100,  NA,  90,  90,  "90",  90,  97,  80)
```
Lets safeguard the code from any other errors that turn numeric into characters

```{r}
x<-student4
x<-as.numeric(x)
x[is.na(x)]=0
mean(x[-which.min(x)])
```
Great! How can we save this function for further use?
We can write our function. 
All functions have at least three things:

-name
-input args
-body

```{r}
grade<-function(x){
x<-as.numeric(x)
x[is.na(x)]=0
mean(x[-which.min(x)])
}
```
play it!

```{r}
grade(student1)
```
Our function works!

## Now grade the whole class
first we have to read the gradebook for the class
```{r}
gradebook<-"https://tinyurl.com/gradeinput"
scores<- read.csv(gradebook, row.names=1)
scores
```


We are going to use the supper useful **apply()**
function to grade all the students with our **grade()** function

```{r}
ans<-apply(scores, 1, grade)
ans
```
## Question 2: Using your grade() function and the supplied gradebook, Who is the top scoring student overall in the gradebook? [3pts]

```{r}
which.max(ans)
```

## Question 3: From your analysis of the gradebook, which homework was toughest on students (i.e. obtained the lowest scores overall? [2pts]

```{r}
ans1<-apply(scores, 2, mean)
ans1
which.min(ans1)

```
have to address the NA values
```{r}
mask <-scores
mask[is.na(mask)]=0
mask

ans2<-apply(mask, 2, mean)
ans2
which.min(ans2)

```
The hardest homework was homework 2.

## Question 4: Optional Extension: From your analysis of the gradebook, which homework was most predictive of overall score (i.e. highest correlation with average grade score)? [1pt]

Here we will use the **cor()** function

```{r}

cor(mask$hw1, ans)
```

```{r}
cor(mask$hw2, ans)
```

```{r}
cor(mask$hw3, ans)
```
I can call the **cor()** function for everything in homework columns and get a value for each, or I can use **apply** and do them all in one go.
```{r}
apply(mask, 2, cor,ans)
```
Homework 5 was the most predictive of overall score. 

Make a boxplot
```{r}
boxplot(scores)
```

#Homework question 6:


